---
title: "Project2_AmyGeisen"
author: "Amy Geisen"
date: "3/10/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Library necessary packages
```{r}
library(klaR)
library(rpart)
library(party)
library(ipred)
library(e1071)
library(ROCR)
library(nnet)
library(MASS)
library(randomForest)
```
#Load MLBench package containing the breast cancer dataset used
```{r}
require(mlbench)
```

#Load Data
```{r}
data(BreastCancer)
# some algorithms don't like missing values, so remove rows with missing values
```

#Clean & Partition Data
#Remove the missing values, create training and evaluation sets by splitting the data 80/20
```{r}
BreastCancer <- na.omit(BreastCancer) 
# remove the unique identifier, which is useless and would confuse the machine learning algorithms
BreastCancer$Id <- NULL 
# partition the data set for 80% training and 20% evaluation (adapted from ?randomForest)
set.seed(2)

ind <- sample(2, nrow(BreastCancer), replace = TRUE, prob=c(0.8, 0.2))
```

#Summarize Data
```{r}
str(BreastCancer)
```

```{r}
summary(BreastCancer)
```

#Create Decision tree model
```{r}
x.rp <- rpart(Class ~ ., data=BreastCancer[ind == 1,])
# predict classes for the evaluation data set
x.rp.pred <- predict(x.rp, type="class", newdata=BreastCancer[ind == 2,])
# score the evaluation data set (extract the probabilities)
x.rp.prob <- predict(x.rp, type="prob", newdata=BreastCancer[ind == 2,])

# To view the decision tree, uncomment this line.
plot(x.rp, main="Decision tree created using rpart")
```
#Create conditional inference tree model
```{r}
x.ct <- ctree(Class ~ ., data=BreastCancer[ind == 1,])
x.ct.pred <- predict(x.ct, newdata=BreastCancer[ind == 2,])
x.ct.prob <-  1- unlist(treeresponse(x.ct, BreastCancer[ind == 2,]), use.names=F)[seq(1,nrow(BreastCancer[ind == 2,])*2,2)]

# To view the decision tree, uncomment this line.
plot(x.ct, main="Decision tree created using condition inference trees")
```
```{r}
x.ct <- rpart(Class ~ ., BreastCancer)
plot(x.ct); text(x.ct) # in "BreastCancer_tree.ps"
summary(x.ct)
x.ct.pred <- predict(x.ct,BreastCancer,type="class")
table(x.ct.pred,BreastCancer$Class)
```
# Create model using random forest and bagging ensemble using conditional inference trees
```{r}
x.cf <- cforest(Class ~ ., data=BreastCancer[ind == 1,], control = cforest_unbiased(mtry = ncol(BreastCancer)-2))
x.cf.pred <- predict(x.cf, newdata=BreastCancer[ind == 2,])
x.cf.prob <-  1- unlist(treeresponse(x.cf, BreastCancer[ind == 2,]), use.names=F)[seq(1,nrow(BreastCancer[ind == 2,])*2,2)]
```

```{r}
x.cf <- randomForest(Class ~ .,BreastCancer)
x.cf.pred <- predict(x.cf, BreastCancer)
table(x.cf.pred, BreastCancer$Class)
```
# Create model using bagging (bootstrap aggregating)
```{r}
x.ip <- bagging(Class ~ ., data=BreastCancer[ind == 1,])
x.ip.prob <- predict(x.ip, type="prob", newdata=BreastCancer[ind == 2,])
```

# Create model using svm (support vector machine)
```{r}
# svm requires tuning
x.svm.tune <- tune(svm, Class~., data = BreastCancer[ind == 1,],
                   ranges = list(gamma = 2^(-8:1), cost = 2^(0:4)),
                   tunecontrol = tune.control(sampling = "fix"))
# display the tuning results (in text format)
x.svm.tune
# If the tuning results are on the margin of the parameters (e.g., gamma = 2^-8), 
# then widen the parameters.
# I manually copied the cost and gamma from console messages above to parameters below.
x.svm <- svm(Class~., data = BreastCancer[ind == 1,], cost=4, gamma=0.0625, probability = TRUE)
x.svm.prob <- predict(x.svm, type="prob", newdata=BreastCancer[ind == 2,], probability = TRUE)
```

```{r}
mysvm <- svm(Class ~ ., BreastCancer)
mysvm.pred <- predict(mysvm, BreastCancer)
table(mysvm.pred,BreastCancer$Class)
```

## plot ROC curves to compare the performance of the individual classifiers


```{r}
png(filename="roc_curve_5_models.png", width=700, height=700)
```

```{r}
# create an ROCR prediction object from rpart() probabilities
x.rp.prob.rocr <- prediction(x.rp.prob[,2], BreastCancer[ind == 2,'Class'])
# prepare an ROCR performance object for ROC curve (tpr=true positive rate, fpr=false positive rate)
x.rp.perf <- performance(x.rp.prob.rocr, "tpr","fpr")
# plot it
plot(x.rp.perf, col=2, main="ROC curves comparing classification performance of five machine learning models")
# Draw a legend.
legend(0.6, 0.6, c('rpart', 'ctree', 'cforest','bagging','svm'), 2:6)

# ctree
x.ct.prob.rocr <- prediction(x.ct.prob, BreastCancer[ind == 2,'Class'])
x.ct.perf <- performance(x.ct.prob.rocr, "tpr","fpr")
# add=TRUE draws on the existing chart 
plot(x.ct.perf, col=3, add=TRUE)

# cforest
x.cf.prob.rocr <- prediction(x.cf.prob, BreastCancer[ind == 2,'Class'])
x.cf.perf <- performance(x.cf.prob.rocr, "tpr","fpr")
plot(x.cf.perf, col=4, add=TRUE)

# bagging
x.ip.prob.rocr <- prediction(x.ip.prob[,2], BreastCancer[ind == 2,'Class'])
x.ip.perf <- performance(x.ip.prob.rocr, "tpr","fpr")
plot(x.ip.perf, col=5, add=TRUE)
# svm
x.svm.prob.rocr <- prediction(attr(x.svm.prob, "probabilities")[,2], BreastCancer[ind == 2,'Class'])
x.svm.perf <- performance(x.svm.prob.rocr, "tpr","fpr")

plot(x.svm.perf, col=6, add=TRUE)
```


```{r}
# Close and save the PNG file.
# dev.off()
```